---
title: "Class part one"
output: html_document
date: "2023-10-13"
---

## R Markdown


```{r}
police <- read.csv("https://raw.githubusercontent.com/asheikhz2/TTU_Zadeh/main/police_app.csv")
head(police)

```
## Lets work with dplyr lybray 

```{r}
data("mtcars")
head(mtcars)
```
## lets import anothery library name fastDummies 
```{r}
library(fastDummies)
mtcars_new <- dummy_cols(mtcars, select_columns = c("cyl","gear"), remove_selected = TRUE, remove_first_dummy = TRUE )
head(mtcars_new)

```

# Matrix Algebra In R

```{r}
A <- matrix(c(1,2,3,4,5,6), ncol=3)
A
```
```{r}
b <-matrix(c(1,2,3,4,5,6,7,8),ncol=4, byrow = T)
b

```
```{r}
b <-matrix(c(1,2,3,4,5,6,7,8),ncol=4)
b
```
```{r}
c <-matrix(c(1,2,3,0,0,0,4,4,3,1,2,5), nrow =4)
c
```
```{r}
d <- matrix(c(1,4,3,"r","c",2,3,4,4), ncol=3)
d

```
```{r}
dim(d)
dim(c)
dim(b)
```
```{r}
nrow(b)
ncol(b)
nrow(d)

```
```{r}
col_vactor= matrix(c(1,2,3,4,5,6), ncol = 1)
col_vactor
```
```{r}
col_vactor[1:3]
d[1:2,1:2]
```
```{r}
a <- matrix(c(1,2,3,2,3,4,1,4,6), ncol = 3)
a
```
```{r}
t(a)

```
```{r}
## regular matrix multiplication (dot product)
dot_product= a %*% a 
dot_product


```
```{r}
#Element wise multipliclation 
em_multipliclation <- a*a
em_multipliclation

```
```{r}
c<- matrix(c(1,2,3,2,3,4), nrow = 3, byrow = T)
c
```
```{r}
el_check <- a %*% c
el_check

```
```{r}
det(a)
```
```{r}
# we use solve function for invers a martix
A <- matrix(c(3,4,5,2,3,4,3,1,4), ncol = 3, byrow = T)
A
```
```{r}
a_invers= solve(A)
a_invers
print("a_invers's determinate")
det(a_invers)
```
```{r}
# inverts multiplicetion 
#inv_mul= A %*% a_invers
inv_mul= A %*% a_invers
round((inv_mul), 10)
```
# inverts multiplicetion gives us identity matrix. 
```{r}
I2= diag(4)
print("I2")

A <- matrix(c(1,2,3,4,5,6,7,8), ncol = 4, byrow = T)
print("A")
print("------")
print(A)

print("------")
print(I2)

```
# so diag function create identity matrix.
```{r}
A %*% I2

```
# discution , if the number of columns and number of rows are same then 
# dot multiplicetion work.
# if we multiply a normal matrix with a identity matrix we will get 
# the result as like as normal matrix
```{r}
I3 <- diag(5)
I3
```
```{r}
a_tranform <-t(A)
print("a_tranform")
a_tranform
print("Multiply with transform matrix")
a_tranform %*% diag(2)


```
```{r}
# lets create a identity matrix 

A <- matrix(c(3,4,5,2,3,4,3,1,4), ncol = 3, byrow = T)
b <- matrix(c(1,4,7,2,5,8,3,6,9), ncol = 3, byrow = T)
b
det(b)
det(A)
A*b
A%*%b
t(b)
solve(A)
I= diag(3)
I
det(I)
solve(I)
I %*% t(A)
I*A
det(I*A)
diag(I*A)
diag(A)

```





